GVN-Hoist: Hoisting Computations from Branches

### 2016-08-19 hoist.tex
exp results

This paper describes code hoisting, a technique that uses the information
computed for PRE to detect identical computations and has a transformation phase
whose goal differs from PRE: it removes identical computations from different
branches of execution.  These identical computations would not be redundant at
run-time and as the number of run-time computations is not reduced, code
hoisting is not a redundancy elimination pass.  The goals of code hoisting are:
\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline by reducing their code size;
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, schedulers can move heavy computations
  earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting expressions out of branches, code hoisting can effectively reduce the
  amount of code to be speculatively executed and can reduce the critical path;
\item to reduce register pressure: by moving computations closer to the
  definitions of their operands;
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted and sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

@@

\section{Experimental Evaluation}
\label{sec:experimental-results}

>>>>

This paper describes code hoisting, a technique that uses the information
computed for PRE to detect identical computations and has a transformation phase
whose goal differs from PRE: it removes identical computations from different
branches of execution.  These identical computations would not be redundant at
run-time and as the number of run-time computations is not reduced, code
hoisting is not a redundancy elimination pass. This is also different from
global code scheduling \cite{dragonbook} in the sense that code-hoisting will
only hoist computations when there are two or more identical computation sharing
a common dominator. The goals of code hoisting are:

\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline by reducing their code size;
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, instruction schedulers can move heavy
  computations earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting expressions out of branches, code hoisting can effectively reduce the
  amount of code to be speculatively executed and can reduce the critical path;
\item to reduce register pressure: by moving computations closer to the
  definitions of their operands;
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted and sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

@@

\section{Experimental Evaluation}
\label{sec:experimental-results}
We ran llvm-testsuite (trunk:d87471f8) with the patch (trunk:86940146 and the
results are listed in Table-\ref{tab:results}. The table lists the number of
scalars, loads, stores and calls hoisted as well as removed. For each category,
the number of instructions removed is greater or equal to the number of
instructions hoisted because each hoisting is performed only when at least one
identical computation is found.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{|l|c|}
      \hline
      Metric                   	     & Number\\\hline
      Scalars hoisted      & 6791  \\\hline
      Scalars removed      & 9696  \\\hline
      Loads hoisted        & 14802 \\\hline
      Loads removed        & 20719 \\\hline
      Stores hoisted       & 15    \\\hline
      Stores removed       & 15    \\\hline
      Calls hoisted        & 8     \\\hline
      Calls removed        & 8     \\\hline
      Total Instructions hoisted & 21616 \\\hline
      Total Instructions removed & 30438 \\\hline
\end{tabular}
  \end{center}
  \caption{Code hoisting metrics on llvm-testsuite}
  \label{tab:results}
\end{table}
